---
title: "S4 classes for representing plates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{S4 classes for representing plates}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(topcorneR)
```

# Wells

A `Well` represents the coordinates of a well in a plate of a given format
(by default, 384).  Given these coordinates, the functions `Row()` and `Column()`
extract the row and column coordinates as letters and numbers respectively.
Note that plate formats names are characters, not numbers.

```{r wells}
well <- Well("C14")
well |> Row()
Column(well)
```

Because the well coordinates come with a plate format, it is possible to
determine where is the "next" well

```{r next_wells}
nextWell(well)
```

# Plates

A plate object represents a physical mutliwell plate that may be filled with
liquids.  In addition to its format, it has a _dead volume_ property that is
used to calculate when it is not possible anymore to draw liquid from a well.

```{r plates}

```

# For developers

## Conventions

 - Class names start with capital letters.
 - Constructor functions have the same name as the class.
