---
title: "S4 classes for representing plates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{S4 classes for representing plates}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(topcorneR) |> suppressPackageStartupMessages()
```

# Wells

A `Well` represents the coordinates of a well in a plate of a given format
(by default, 384).  Given these coordinates, the functions `Row()` and `Column()`
extract the row and column coordinates as letters and numbers respectively.
Note that plate formats names are characters, not numbers.

```{r wells}
well <- Well("C14")
well |> Row()
Column(well)
Column(Well("C04"))
```

Because the well coordinates come with a plate format, it is possible to
determine where is the "next" well.

```{r next_wells}
nextWell(Well("A24"))
```

# Plates

A plate object represents a physical mutliwell plate that may be filled with
liquids.  In addition to its format, it has a _dead volume_ property that is
used to calculate when it is not possible anymore to draw liquid from a well,
and a _maximal volume_ property that is used to to calculate when it is not
possible anymore to add liquid to a well.  These properties can be left
unspecified when the plate representation is for theoretical purposes only.  All
volumes are expressed in nanoliters.

```{r plates}
(plate <- Plate("96", deadVolume = 1e4, maxVolume = 1e5))
Plate("384")
```

The function `PlateTypeToWellNames` enumerates the names of the wells of a
plate.

```{r plate_functions}
PlateTypeToWellNames("96") |> head(13)
```

A filter function, `setWell`, is provided to add contents to a plate.

```{r set_well}
plate <- plate |>
  setWell(Well("A01"), "dNTP", 1e5) |>
  setWell(Well("A02"), "DMSO", 1e5)
```

The contents of a plate can be extracted by converting _Plate_ objects to their
base class, _DataFrame_.

```{r plate_contents}
plate |> DataFrame() |> head()
```

Note that setting again the same reagent overrides the volume, and that setting
another reagent does not reset the previous one.

```{r set_well2}
plate |> setWell(Well("A01"), "dNTP", 1e4) |> DataFrame() |> head(3)
plate |> setWell(Well("A01"), "DMSO", 1e4) |> DataFrame() |> head(3)
```

The `sourceReagent` function reports what reagents are found in the whole plate
or one of its wells.

```{r source_reagent}
sourceReagent(plate)
sourceReagent(plate, Well("A02"))
```

Conversely, `plateWellVolume` reports the volume of a well or of one of its
reagents.

```{r plate_well_volume}
plate <- plate |>
  setWell(Well("A03"), "DMSO", 1e4) |>
  setWell(Well("A03"), "H2O", 1e4)
plateWellVolume(plate, Well("A03"))
plateWellVolume(plate, Well("A03"), "DMSO")
```

Finally, starting from `A01` or an arbitrary location, it is possible to find
where is the next well containing a given reagent.

```{r plate_seek_reagent}
seekReagent(plate, "H2O")
```

# For developers

## Conventions

 - Class names start with capital letters.
 - Constructor functions have the same name as the class.
